---
title: "Presentation"
author: "s210136, s200177, s202754"
date: "5/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Define functions -------------------------------------------------------------

```{r}
source(file = "R/99_project_functions.R")
```


Initial loading of the data

```{r}

```











Cleaning the data:
Load data --------------------------------------------------------------------

```{r}
# Patient 1
patient_1 <- read_csv(file = "Data/01_patient_1I.csv.gz")

# Patient_2
patient_2 <- read_csv(file = "Data/01_patient_2C.csv.gz")

# Patient_3
patient_3 <- read_csv(file = "Data/01_patient_3C.csv.gz")

# Patient_4
patient_4 <- read_csv(file = "Data/01_patient_4CO.csv.gz")

# Patient_5
patient_5 <- read_csv(file = "Data/01_patient_5CO.csv.gz")

# Patient_6
patient_6 <- read_csv(file = "Data/01_patient_6I.csv.gz")

```


Combining patients -----------------------------------------------------------

Wrangle data -----------------------------------------------------------------

 ------------------- Getting gene names and cell bar codes -------------------

Loading metadata and getting cell types ------------------------------------
Metadata supplied by the authors of the paper("ids.csv")

```{r}
gene_mapping <- read_csv("Data/_raw/ids.csv",
                         col_names=c("value","Gene")) %>%
  tibble()

```

Transpose so that genes are columns and rows are cells

```{r}
patient_1 <- patient_1 %>% 
  rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from = rowname,
              values_from = value)

Genes <- read_csv("Data/Genes_1")
Genes <- Genes %>%
  left_join(gene_mapping) %>%
  pluck("Gene")

```


Rename genes and make nice name for column with barcodes

```{r}
colnames(patient_1) <- c("Cell_Barcode",
                         Genes)

```


-------------- now do the same for 5 other patients --------------


```{r}
# Patient_2
patient_2 <- patient_2 %>%
  rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from = rowname,
              values_from = value)

Genes <- read_csv("Data/Genes_2")
Genes <- Genes %>% 
  left_join(gene_mapping) %>% 
  pluck("Gene")

colnames(patient_2) <- c("Cell_Barcode",
                         Genes)
# Patient_3
patient_3 <- patient_3 %>% 
  rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from = rowname,
              values_from = value)

Genes <- read_csv("Data/Genes_3")
Genes <-Genes %>% 
  left_join(gene_mapping) %>% 
  pluck("Gene")

colnames(patient_3) <- c("Cell_Barcode",
                         Genes)

#Patient_4               
patient_4 <- patient_4 %>% 
  rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from = rowname,
              values_from = value)

Genes <- read_csv("Data/Genes_4")
Genes <- Genes %>% 
  left_join(gene_mapping) %>% 
  pluck("Gene")

colnames(patient_4) <- c("Cell_Barcode",
                         Genes)

#Patient_5               
patient_5 <- patient_5 %>% 
  rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from = rowname,
              values_from = value)

Genes <- read_csv("Data/Genes_5")
Genes <- Genes %>% 
  left_join(gene_mapping) %>% 
  pluck("Gene")

colnames(patient_5) <- c("Cell_Barcode",
                         Genes)

#Patient_6                
patient_6 <- patient_6 %>% 
  rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from = rowname, 
              values_from = value)

Genes <- read_csv("Data/Genes_6")
Genes <- Genes %>% 
  left_join(gene_mapping) %>% 
  pluck("Gene")

colnames(patient_6) <- c("Cell_Barcode",
                         Genes)

```




 Write data -------------------------------------------------------------------
 
 
```{r}
# Patient_1
write.csv(patient_1,
          file = "Data/02_patient_1I.csv.gz")

# Patient_2
write.csv(patient_2,
          file = "Data/02_patient_2C.csv.gz")

# Patient_3
write.csv(patient_3,
          file = "Data/02_patient_3C.csv.gz")

# Patient_4
write.csv(patient_4,
          file = "Data/02_patient_4CO.csv.gz")

# Patient_5
write.csv(patient_5,
          file = "Data/02_patient_5CO.csv.gz")

# Patient_6
write.csv(patient_6,
          file = "Data/02_patient_6I.csv.gz")
```
 
 


Remove Data ------------------------------------------------------------------

```{r}
rm(patient_1,
   patient_2,
   patient_3,
   patient_4,
   patient_5,
   patient_6,
   Genes,
   gene_mapping)

```

